Normal States Training for High Accuracy Event Detection

### 2016-07-08 paper/ieeeFormat.tex
before modifying

\section{Experiments}
\subsection{Effects of Normal States Initialization}
We use 20newsgroups\cite{lang1995newsweeder} dataset\footnote{http://qwone.com/~jason/20Newsgroups} to check the effects of normal states initialization.
\begin{table}[]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{|l|l|}
\hline
comp.graphics            & Category:Computer graphics      \\ \hline
comp.os.ms-windows.misc  & Category:Microsoft Windows      \\ \hline
comp.sys.ibm.pc.hardware & Category:IBM personal computers \\ \hline
comp.sys.mac.hardware    & Category:Macintosh computers    \\ \hline
comp.windows.x           & Category:X Window System        \\ \hline
rec.autos                & Category:Automobiles            \\ \hline
rec.motorcycles          & Category:Motorcycles            \\ \hline
rec.sport.baseball       & Category:Baseball               \\ \hline
rec.sport.hockey         & Category:Hockey                 \\ \hline
sci.crypt                & Category:Cryptography           \\ \hline
sci.electronics          & Category:Electronics            \\ \hline
sci.med                  & Category:Medicine               \\ \hline
sci.space                & Category:Outer space            \\ \hline
misc.forsale             & Category:Sales                  \\ \hline
talk.politics.misc       & Category:Politics               \\ \hline
talk.politics.guns       & Category:Gun politics          \\ \hline
talk.politics.mideast    & Category:Middle East           \\ \hline
talk.religion.misc       & Category:Religion               \\ \hline
alt.atheism              & Category:Atheism                \\ \hline
soc.religion.christian   & Category:Christians             \\ \hline
\end{tabular}
\end{table}

@@

@@

Here we present the effectiveness of our proposed algorithm UMIETM and the efficiency of its online performance.
We evaluate the efficiency by perplexity, precision for event detection. 
We check the time cost and complete likelihood for efficiency.

To improve the quality of analyzing on tweets, we do necessary pre-processing: (1) splitting dataset by week, (2) segmenting Chinese words, (3) removing stop words and low frequency words whose document frequency in its time window is less than 3.

We compare our model UMIETM with twitterLDA, timeUserLDA, Author-LDA, and our model's variant IETM.
Author-LDA combines the tweets posted by same author into a single document, then run standard LDA on the assembled tweets.
TwitterLDA is designed for topic modeling on twitter.
We compare with Author-LDA and TwitterLDA for confirming the significance of distinguishing user interests from events in microblog.
TimeUserLDA is designed for retrospective event detection in microblog, and considers to distinguish events from user interests.
We compare with timeUserLDA to show the impact of user profile.
The variant IETM (\textit{Interest and Event Topic Model}) models user's interests and events without the help of user profiles.

We set the asymmetric \(\alpha\) and symmetric \(\beta=0.01\) for UMIETM, where \(\alpha\) will be optimized by Gibbs EM algorithm\cite{wallach2008structured}.
\(\alpha=0.1\), \(\beta=0.01\) are set for all remaining models. 
After cross validation we find that UMIETM and IETM perform best on \(K=90\) and \(E=30\), \(\kappa=0.01\).
To compare equally, we set the same topic number for Author-LDA, twitterLDA, timeUserLDA (which means when we run models offline on 1 time window, we shall set 120 topics for all, and 150 topics on 2 time windows and so on).

To verify the role of user profiles played, we set UMIETM(-) as the degradation of UMIETM, which take symmetric prior \(\alpha\).

In this subsection, we illustrate the performance of UMIETM in which user profile is considered.

\textbf{Quantitative Measure.}
We initialize UMIETM and IETM by batch learning on data from first week to third week, then run them in online learning way from fourth week to ninth week.
On each week we calculate their perplexities\cite{wallach2009evaluation}, where \(perplexity(D_{test})=\exp{\{-\frac{\sum_{u=1}^{U}\sum_{d=1}^{D_u}\log{p(w_{ud})}}{\sum_{u=1}^{U}\sum_{d=1}^{D_u}N_{ud}}\}}\) and \(p(w_{ud})=(1-\pi_u)\sum_{k=1}^{K}\theta_{uk}\prod_{n=1}^{N_{ud}}(\phi_{s,w_{udn}}(1-\rho)+\phi_{k,w_{udn}}\rho)+\pi_u \sum_{k=1}^{K}\eta_{t,k}\prod_{n=1}^{N_{ud}}(\phi_{s,w_{udn}}(1-\rho)+\phi_{k,w_{udn}}\rho)\).
The others are trained from first week to third week, and the held out perplexities are calculated on data from fourth week to ninth week.
TimeUserLDA and IETM's perplexities are smaller than User-LDA, twitterLDA as they both consider distinguishing user interests from events.

In Table \ref{tab:heldoutPerplexity} the perplexity of UMIETM is 3107.8, and much smaller than others. It demonstrates that user profile is significant for tweet stream's modeling.

In Table \ref{tab:metrics}, we evaluate the events detected by models. We asked the annotators to label the event with score 1, and non-event related topic as 0. 
The precision and recall is illustrated, where UMIETM(-) performs slightly better than IETM. 
And UMIETM outperforms UMIETM(-) in detecting high quality event. 
A reasonable analysis is that UMIETM uses the profile information sufficiently by asymmetric priors. 
In this way, we also prove that the well exploiting of user profile information is important to model user interests and events on tweets. 
%\subsubsection{Precision of Event Detection}

\textbf{Case Study} 
Some events detected by our UMIETM model are shown in Table \ref{fig:event}.
Comparing with UMIETM, timeUserLDA fails to discover the \textit{shoddy construction} event in the second week, while IETM reports this event as \textit{bi, women, elegant, adoption, engineering, reed}. Obviously IETM fails to distinguish this event from user interests.
UMIETM filters users' interests like \textit{bi, women, elegant, adoption} using their profile {\#baby, \#women}, and detect the event.
\subsection{Efficiency}
We implement our methods on mallet\footnote{http://mallet.cs.umass.edu/dist/mallet-2.0.7.tar.gz}, and run them on Linux server with 8 cores(2.00GHz) and 64GB memory.
In this subsection, we verify the convergence and online performance of our online learning algorithm.

\textbf{Convergence.}
The convergence of algorithm is vital in real time streaming environment. 
UMIETM needs several Gibbs sweeps to find out the optimal parameters of model.
But the size of data is usually large even spitted into time windows.
The time complexity of UMIETM is \(O(I_1K|P|+I_2K|W|+I_2E|W|)\) given in section \ref{subsec:online}.
It suggests that the more iterations Gibbs sampling have to do, the less efficiently our model performs. 
Fortunately UMIETM is very economic to its stable state and does not need so much iterations.
As suggested by \cite{wallach2009rethinking} and \cite{jian2014factorOfTopicModel}, we choose LDA with asymmetric priors as baseline, which is much more comparable than standard LDA on microblog dataset.
We run UMIETM on the first three time windows. 
Correspondingly we only use the tweets in these time windows for LDA with asymmetric priors.
Burn-in period is set to 20 and optimize interval also 20, which means the priors are optimized every 20 iterations as the red curve in Fig \ref{fig:completeLoglikelihood}.

One of the criteria for stopping Gibbs sampling is that the complete log likelihood is stabilized.
We stop UMIETM after 50 iterations since the complete log likelihood only increases 0.02\% in last iteration.
It will take LDA(asymmetric priors) 300 minutes to run 1000 iterations, and 20 minutes per 10 iterations for UMIETM.
There is a trade-off between the effectiveness by more iterations and efficiency by less,  we set the iteration number as 10 in online setting.
After 10 iterations the UMIETM's complete log likelihood will increase no more than 0.2\%.

\textbf{Online Performance.}
As verified in the above subsection \textit{Convergence}, we run 10 Gibbs sweeps for each time window in online learning phase.
When UMIETM is implemented in online way, it doesn't have to revisit the tweets in previous windows. 
So its time cost in each window is proportional to the data size of current window  as shown in Figure \ref{fig:onlinelearningResult}.
In average, online UMIETM can process five thousand tokens per second.
%Memory cost grows slightly and the increased memory cost only depends on the size of vocabulary.

LSH based event detection on microblogs is used by \cite{petrovic2010streaming},\cite{mccreadiescalable}. 
The former one detects events as the cluster of tweets by LSH, and the latter one extends this method on distributed machines. 
They both only use the cosine similarity without considering semantic meaning of words. 
So LSH based method may split the same event into clusters of similar tweets.
However LSH based method reports 10578 events in 7th time window, and 10901 events in 8th window. 
The other problem of LSH based event detection method is that the bad design may lead unstable performance. 
The tweets are mapped into LSH defined space in a skew way which is illustrated in Fig \ref{fig:onlinelearningResult}.
The other comparison is timeUserLDA which is designed for retrospective event detction. 
The purple line in Fig \ref{fig:onlinelearningResult} goes up straightly as timeUserLDA has to revisit previous tweets to make a decision whether the tweet is event related. 
The perplexity of online UMIETM on the 11 time window in Fig \ref{fig:onlinelearningResult} is \(3036.44\pm397.14\). 
And the average precision of detecting high quality event is 0.22.
Both indicators demonstrate that UMIETM can perform well in online learning way. 
\subsection{Application on Reddit}

>>>>

\section{Experiments}
\subsection{Effects of Normal States Initialization}
We use 20newsgroups\cite{lang1995newsweeder} dataset\footnote{http://qwone.com/~jason/20Newsgroups} to check the effects of normal states initialization.

What's NPMI? maybe can refer to \cite{Aletras2013EvaluatingTC}\cite{Rder2015ExploringTS}\footnote{https://github.com/AKSW/Palmetto}.
\begin{table}[]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{|l|l|}
\hline
comp.graphics            & Category:Computer graphics      \\ \hline
comp.os.ms-windows.misc  & Category:Microsoft Windows      \\ \hline
comp.sys.ibm.pc.hardware & Category:IBM personal computers \\ \hline
comp.sys.mac.hardware    & Category:Macintosh computers    \\ \hline
comp.windows.x           & Category:X Window System        \\ \hline
rec.autos                & Category:Automobiles            \\ \hline
rec.motorcycles          & Category:Motorcycles            \\ \hline
rec.sport.baseball       & Category:Baseball               \\ \hline
rec.sport.hockey         & Category:Hockey                 \\ \hline
sci.crypt                & Category:Cryptography           \\ \hline
sci.electronics          & Category:Electronics            \\ \hline
sci.med                  & Category:Medicine               \\ \hline
sci.space                & Category:Outer space            \\ \hline
misc.forsale             & Category:Sales                  \\ \hline
talk.politics.misc       & Category:Politics               \\ \hline
talk.politics.guns       & Category:Gun politics          \\ \hline
talk.politics.mideast    & Category:Middle East           \\ \hline
talk.religion.misc       & Category:Religion               \\ \hline
alt.atheism              & Category:Atheism                \\ \hline
soc.religion.christian   & Category:Christians             \\ \hline
\end{tabular}
\end{table}

@@

\begin{figure}
        \centering
        \includegraphics[width=1.0\columnwidth]{img/NPMI.pdf}
        \caption{NPMI}
\end{figure}

@@

\subsection{Application on Reddit}
